---
title: "p8105_hw2_ww1234"
author: "Jason Wang"
date: "10/2/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(tidyr)
library(dplyr)
library(lubridate)


```

# problem 1
##load and clean data for Mr. Trash Wheel
```{r, question1, echo = TRUE}
## load in a dataset

trash_wheel_data = read_excel(path = "../data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
                              sheet = "Mr. Trash Wheel",
                              range = cell_cols("A:N")
                              ) %>% 
  janitor::clean_names() %>% 
  na.omit() 
trash_wheel_data$sports_balls = round(as.integer(trash_wheel_data$sports_balls), digits = 0)

###checkpoints###
head(trash_wheel_data,5)
summary(trash_wheel_data)
count(trash_wheel_data)
###checkend###

```

##Read and clean precipitation data of 2017

```{r, echo = TRUE}
##read and clean precipitation data of 2017
precipitation_dataset_2017 = read_excel(path = "../data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                                        sheet = "2017 Precipitation",
                                        range = "A2:B14") %>% 
  na.omit() %>%
  janitor::clean_names() %>% 
mutate( year = '2017' )
 
####check points####
view(precipitation_dataset_2017)
summary(precipitation_dataset_2017)
count(precipitation_dataset_2017)
####checkend#####

```


##read and clean precipitation data of 2018
```{r, echo = TRUE}
precipitation_dataset_2018 = read_excel(path = "../data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                                        sheet = "2018 Precipitation",
                                        range = "A2:B14") %>% 
  na.omit() %>% 
  janitor::clean_names() %>%
mutate( year = '2018' )

####check points####
view(precipitation_dataset_2018)
summary(precipitation_dataset_2018)
####checkend#####

```

##combination the two dataset of precipitation_dataset_2017 and precipitation_dataset_2018
```{r, echo = FALSE}
precipitation_dataset = left_join(
  precipitation_dataset_2017, 
  precipitation_dataset_2018, 
  by = "month",
  suffix = c("_2017", "_2018")
) %>% 
  select("month", "total_2017", "total_2018") %>% 
  rename("2017" = "total_2017", "2018" = "total_2018") 
precipitation_dataset$month = month.name[precipitation_dataset$month]

######checkpoints####
view(precipitation_dataset)
summary(precipitation_dataset)
######checkend####


```

the number of observation in 2017 is `r count(precipitation_dataset_2017)`.
the number of observation in 2018 is `r count(precipitation_dataset_2018)`.
the total precipitation in 2018 is `r sum(precipitation_dataset_2018$total, na.rm = TRUE)`.


Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2017?

# problem 2
```{r, pols, echo = TRUE}
##load and clean data in pols-month.csv
pols_month_dataset = read_csv(file = "../data/pols-month.csv") %>%   ##load pols-month.csv
  janitor::clean_names() %>% ##clean the dataset
  separate(col = "mon", sep = "-", into = c("year", "month", "day"), 
           convert = type.convert("month", "year", "day" , 
                                  numerals = c("allow.loss"))) %>%  ##separate to year, month, day, and transfer month format "01" "02"... to "1", "2)
  transform(month = month.abb[month]) %>%   ##replace month number with month name
  mutate(president = case_when(
    prez_dem == 1 ~ "dem",
    prez_gop == 1 ~ "gop",
    TRUE ~ ""
  )) %>%  ##add variable president with value "dem" and "gop"
  select("year", "month", "gov_gop":"rep_gop", "gov_dem":"rep_dem", "president") ##remove prez_gop, prez_dem and day variable

summary(pols_month_dataset)
count(pols_month_dataset)


```

```{r, snp, echo = TRUE}
##load and clean data in snp.csv
snp_dataset = read_csv(file = "../data/snp.csv") %>% ##load the data
  janitor::clean_names() %>%                         ##clean the data
  separate(col = "date", 
           sep = "/", 
           into = c("month", "day", "year"), 
           convert = type.convert("month", "day", "year", 
                                  numerals = c("allow.loss"))) %>% 
  transform(month = month.abb[month]) %>%            ##separate the data
  select("year", "month", "day", everything()) 

```

```{r, unemployment, echo = TRUE}
##load and clean unemployment data
unemployment_dataset = read_csv(file = "../data/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
  ) %>% 
  janitor::clean_names()
  

```

```{r, merg_snp_pols_unem, echo = TRUE}
## first merge the snp and pols dataset
merg_snp_pols_dataset = full_join(snp_dataset, pols_month_dataset, by = c("year", "month"))

merg_snp_pols_unem_dataset = left_join(merg_snp_pols_dataset, unemployment_dataset, by = c("year", "month"))

```

summary
Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables). 

# problem 3
```{r, babyname, echo = TRUE}
#load and clean the data
popular_baby_name_dataset = read_csv(file = "../data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  distinct()

popular_male_baby_name_dataset = popular_baby_name_dataset %>% 
  filter(gender == "MALE") %>% 
  filter(ethnicity == "WHITE NON HISPANIC") %>% 
  filter(year_of_birth == 2016)
 
plot(popular_male_baby_name_dataset$rank, popular_male_baby_name_dataset$count,
     xlab = "the number of children with a name",
     ylab = "the rank in popularity of that name")
  
```


